---
title: "Lab2"
author: "Jonathan Douglas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Intro2R)
```

# Tasks:

## Task 1: Download Files and Set Working Directory
```{r}
setwd("/Users/jonathandouglas/MATH4753/FALL244753doug0080/Labs/Lab2")

#Check the working directory
getwd()
```

## Task 2: Change file type and read in data
```{r}
#Read in the datafile
EPAGAS = read.csv("EPAGAS.csv")

#Print the first 6 lines
head(EPAGAS)
```

## Task 3:  Use the emperical formula to find outliers of mpg
```{r}
#Create mpg vector
mpg = EPAGAS$MPG

#Use the emperical formula for z
z = (mpg - mean(mpg)) / sd(mpg)

#Verify the results 
print("Results for verification:")
#Mean of std deviations
mean(z) 
print("(essentially 0)")
#standard deviation of std deviations
sd(z)**2

#Use the [] methods to subset data that can be considered an outlier or possible outlier
print("Possible outliers:")
mpg[abs(z) >= 2.0 & abs(z) <= 3.0]
```


```{r}
print("Statistical outliers:")
mpg[abs(z) > 3.0]
```

```{r}
#Use the lattice package to make the boxplot with colored dots
#First, sort the dots into the different colors
# Plot the outliers in red and the possible outliers in blue
outlier_col = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))
```

```{r}
#Now, create the dotplot
library(lattice)
dotplot(mpg,col=outlier_col)
```

## Task 4: 
```{r}
ggplot(data.frame(mpg), aes(x = mpg)) + geom_boxplot(notch = TRUE, notchwidth = 0.5, orientation = 'y', staplewidth = 0.35, fill = "black") + ggtitle("Boxplot of EPA MPG")
#Create the scatter plot
#ggplot(ddt, aes(x = WEIGHT, y = LENGTH, color = SPECIES)) + geom_point()+ ggtitle("Jonathan Douglas")

```

#### Step 2: Use Chebyshev’s theorem
```{r}
#Predict the portion of the data that within two standard deviations of the mean.
#Use the formula 1 - (1/z^2):
print("Portion of data within two std deviations")
1 - (1/2**2)
```

#### Step 3: Use R to caluclate the proportion within 2 std deviations of the mean
```{r}
#Find the length of data within two std deviations
num_in_two = length(mpg[abs(z) < 2])
#Find the length of mpg 
total_length = length(mpg)

#Find the percentage of data in 2 std deviations
print("Fraction of data in two standard deviations:")
num_in_two / total_length
```

Chebushev's Theorem does support this data.

#### Step 4: Use the empirical rule to find the fraction of data in 2 std deviations

The emperical rule finds that approximately 95% of data is found in the second standard deviation.
This corresponds very well to the observation that 96% of the data is in the 2 standard deviations.

Can the empirical formula be valid here?
Check for:
1. Unimodality
2. Symmetric through the mode

```{r}
hist(mpg)
```
This dataset exhibits these conditions, so the Emperical rule is valid in this case.

