---
title: "Lab#5"
author: "Jonathan Douglas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1: Set and get the working directory

Set and Find the WD

```{r}
setwd("/Users/jonathandouglas/MATH4753/FALL244753doug0080/Labs/Lab5")
getwd()
```

## Task 2: Create sample distributions
```{r}
## sampling function
# iter = iterations, n=sample size
# set default values
mybin=function(iter=100,n=10, p=0.5){ 
  # make a matrix to hold the samples
  #initially filled with NA's
  sam.mat=matrix(NA, nr=n, nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes in each trial
  succ=c()
  for( i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(c(1,0), n, replace=TRUE, prob=c(p,1-p))
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  #Make a table of successes
  succ.tab=table(factor(succ, levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
  succ.tab/iter
}

```

#### Distribution with 100 iterations
```{r}
#100 iterations
mybin(iter=100,n=10, p=0.7)
```

####Distribution with 200 iterations
```{r}
#200 iterations
mybin(iter=200,n=10, p=0.7)
```

#### Distribution with 500 iterations
```{r}
#500 iterations
mybin(iter=500,n=10, p=0.7)
```

#### Distribution with 1000 iterations
```{r}
#1000 iterations
mybin(iter=1000,n=10, p=0.7)
```

#### Distribution with 10000 iterations
```{r}
#10000 iterations
mybin(iter=10000,n=10, p=0.7)
```

#### Check this graph with the binomial distribution
```{r}
#dbinom function
dbinom(0:10, size = 10, prob = 0.7)
```

This distribution matches very strongly to the table from the last distribution.

## Task 3: Bag of marbles problem

####Sample without replacement
n = 20, p = 12/20
```{r}
#Use sample function for the marble problem.
#First, we are going to create a vector of 1s and 0s
marble_color = c(rep(1, 12), rep(0, 8))
sample(marble_color, size = 5, replace = FALSE)
```

#### Sample with replacement
```{r}
#Use sample function for the marble problem.
#First, we are going to create a vector of 1s and 0s
marble_color = c(rep(1, 12), rep(0, 8))
sample(marble_color, size = 5, replace = TRUE)
```

#### Use the myhyper() function
```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
  # make a matrix to hold the samples
  #initially filled with NA's
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes over the trials
  succ=c()
  for( i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
  succ.tab/iter
}
```

#### myhyper call with 100 iterations
```{r}
#myhyper call with 100 iterations
myhyper(iter=100,n=5, N=20,r=12)
```

#### myhyper call with 200 iterations
```{r}
#myhyper call with 200 iterations
myhyper(iter=200, n=5, N=20, r=12)
```

#### myhyper call with 500 iterations
```{r}
#myhyper call with 500 iterations
myhyper(iter=500,n=5, N=20,r=12)
```

#### myhyper call with 1000 iterations
```{r}
#myhyper call with 1000 iterations
myhyper(iter=1000,n=5, N=20,r=12)
```

#### myhyper call with 10000 iterations
```{r}
#myhyper call with 10000 iterations
myhyper(iter=10000,n=5, N=20,r=12)
```

#### Verify with the dhyper() function
```{r}
dhyper(x=0:5, m=12, n=8, k=5)
```

Our hyper function matches farily consistently with the R dhyper function

## Task 4: mysample function

#### Import the mysample function
```{r}
#The mysample function
mysample=function(n, iter=10,time=0.5){
  for (i in 1:iter){
    #make a sample
    s=sample(1:10,n,replace=TRUE)
    # turn the sample into a factor
    sf=factor(s,levels=1:10)
    #make a barplot
    barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
    main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
    ylim=c(0,0.2)
    )
    
    #release the table
    Sys.sleep(time)
  }
}

#mysample(n=1000, iter=30, time = 1)#This is commented out
mysample(n=1000, iter=1, time = 1)#This is ran so that only one plot appears
```

What this function is doing is it creates random samples of numbers 1 -10 and does this random sample n number of times. It then compiles the fraction of each sampled value into a bar chart, and the chart is printed. This process is repeated iter number of times. 

## Task 5: Sample Calculations

#### Calculate the combinations (choose function)
```{r}
#Calculate combinations
x1 = choose(n = 8, k = 4)
cat("8 choose 4:", x1)
```

#### Poisson Distribution
```{r}
#Find the probability of a poisson distribiution for y ≤ 4
prob_less_4 = ppois(4, lambda = 2)

#Find probability of greater than 4
cat("Probability of y > 4:", (1 - prob_less_4))
```

#### Negative Binomial Distribution
```{r}
#  This means that in R we need n=3 (size), x=y-r=10-3=7
r_nbinom = dnbinom(7,3,0.4)  # Nu Failures, Nu of successes, prob success
mynbin=function(y,r,p){
  choose(y-1,r-1)*p^r*(1-p)^(y-r)
}

cat("R Nbinom function:", r_nbinom, '\n')
cat("Our Negative Binom Function:", mynbin(10,3,0.4))

```

#### Binomial Distribution
```{r}
#Find the probability of P <= 8
cat("P <= 8:", pbinom(8, 15, 0.4))
```

## Task 6: Test my new function

This is the function mybin
```{r}
#Import my package
library(FALL244753doug0080)

#Run the function mybin from above
FALL244753doug0080::mybin(iter=235,n=15, p=0.4)#Here is a test run of the new function
```

This function works as expected with no problems.








