---
title: "Lab1"
author: "Jonathan Douglas"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

# Tasks

## Task 1: Set the Working Directory

```{r}
setwd("/Users/jonathandouglas/MATH4753/FALL244753doug0080/Labs/Lab1")
getwd()
#Hello
```

## Task 2: Read in the DDT File as a CSV
```{r}
ddt <- read.csv("DDT-1.csv", header = TRUE)
head(ddt)
```

## Task 3: DDT Data manipulation

#### Step 1: Sort Variables as Qualitative and Quantitative
```{r}
#ddt already exists, Print the head of ddt
head(ddt)
```
The qualitative variables in dtt are "River" and "Species". The quantitative variables are mile, length, weight, and ddt.

#### Step 2: To find the number of species: Use the unique() method
```{r}
#Number of species in the dataset
unique(ddt$SPECIES) %>% length()
```

#### Step 3: Subset the dataset so to show the observations with SPECIES of LMBASS and WEIGHT bigger than 800gms
```{r}
#Filter and print the data
ddt %>% filter(SPECIES == "LMBASS" & WEIGHT > 800)
```

#### Step 4: Subset the dataset so to show the observations where the RIVER is SCM and DDT bigger than 4.0

```{r}
#Filter and print the data
ddt %>% filter(RIVER == "SCM" & DDT > 4.0)
```

## Task 4: Clicker Questions

#### Question 1: The mean length of fish in Dataframe

```{r}
#Calculate the mean length of the fish
mean(ddt$LENGTH)
```

#### Question 2: Calculate standard deviation of the weight of the fish
```{r}
#Standard deviation from weight
sd(ddt$WEIGHT)
```
#### Question 3: Plot Length vs Weight
```{r}
#Plot Length vs Weight
plot(LENGTH~WEIGHT, data = ddt)
```

False: The graph above is not the one seen in the lab, as the axis are flipped. (0)

#### Question 4: What does v/20 return
```{r}
v = 1:20
v/20
```

The last value is 1.00

## Task 5: Chart Creation
#### Step 1: Make a table of the rivers
```{r}
#Create a table of the river
riv=with(ddt,table(RIVER))
riv
```

#### Step 2: Make a colored barplot of the rivers
```{r}
#Create a color-coded barplot of the rivers
barplot(riv,col=rainbow(4))
```

#### Step 3: Create a table of rivers crossed with the species of fish
```{r}
#Create a table of the river
riv_sp=with(ddt,table(RIVER, SPECIES))
riv_sp
```
#### Step 4: Create a barplot with rivers crossed with species of fish
```{r}
#Create a color-coded barplot of the rivers
barplot(riv_sp,beside = TRUE, col=rainbow(4))
```

## Task 6: Pie Charts
#### Step 1: Make a pie chart of fish species
```{r}
#Plot pie chart of fish species
sp = with(ddt, table(SPECIES))
pie(sp, col = 1:4)
```

#### Step 2: Make a pie chart of the rivers
```{r}
#Plot pie chart of fish species
sp = with(ddt, table(RIVER))
pie(sp, col = 1:4)
```

## Task 7: Box Plots

#### Step 1: Create a boxplot of the variable DDT
```{r}
#Create a boxplot of the DDT
boxplot(ddt$DDT)
```

#### Step 2: Make a boxplot of WEIGHT
```{r}
#Plot a boxplot of WEIGHT
boxplot(ddt$WEIGHT)
```

#### Step 3: Plot the boxplot of the variable LENGTH
```{r}
#Create boxplot of length
boxplot(ddt$LENGTH)
```

## Task 8: Coplots
#### Step 1:	Make a coplot of LENGTH Vs WEIGHT given RIVER
```{r}
#embellish the plot with species using color using 'ifelse(test,yes,no)'
fishcol=with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))

coplot(LENGTH~WEIGHT|RIVER,data=ddt,col=fishcol)
```

#### Step 2: Make a coplot of DDT Vs WEIGHT given SPECIES
```{r}
coplot(DDT~WEIGHT|SPECIES,data=ddt,col=fishcol)
```

## Task 9: Recreate the plots is the lab

#### Step 1: Boxplot
```{r}
#Create boxplot with our name on it using ggplot
ggplot(ddt, aes(x = SPECIES, y = WEIGHT, fill = RIVER)) + geom_boxplot() + ggtitle("Jonathan Douglas")
```

#### Step 2: Create the second plot: Violin plot
```{r}
#Create the violin plot
ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) + geom_violin() + ggtitle("Jonathan Douglas")
```

#### Step 3: Create the scatterplot
```{r}
#Create the scatter plot
ggplot(ddt, aes(x = WEIGHT, y = LENGTH, color = SPECIES)) + geom_point()+ ggtitle("Jonathan Douglas")
```

