---
title: "Assignment1"
author: "Jonathan Douglas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Intro2R)
setwd("/Users/jonathandouglas/MATH4753/FALL244753doug0080/Assignments/Assignment1/")
```

# Questions

## Question 1: How will grading be done this semester?

Grading will be weighted in the following distribution:
15% - The four assignments
10% - Labs
10% - Projects: 1/3 of grade on project 1, 2/3 on project 2
10% - In-class quizzes
5% - Chapter quizzes
20% - Midterm Exams
30% - Final

Grades Ranges:
A - 90s
B - 80s
C - 70s and 60s
D - 50s
F - <50%

## Question 2: 

## Question 3:

a. Quantitative
b. Quantitative
c. Qualitative
d. Quantitative
e. Qualitative
f. Quantitative
g. Qualitative

## Question 4:Types of Sampling

#### Simple Random Sampling:
A simple random sample ensures that every group of items in a population that is sampled has an equal chance of selection.

#### Stratified Random Sampling:
This design is used when the sampling population can be seperated into groups and shares more similarities with other members in each group than across different groups. Representative samples from each group (strata) are included in the sample. 

#### Cluster Sampling:
If data in a population is already in natural groups, then it can sometimes be more convenient to take a random sample of the larger groups, then look at the population within.

#### Systematic Sampling:
This method of sampling involves selecting every nth item in a population to be sampled. 

## Question 5: 1.15 - Well Problem
```{r}
#Sample code
mtbe=read.csv("MTBE.csv", header=TRUE) # You will need to change the address
#Find sample of 5 wells 
ind = sample(1:223, 5, replace=FALSE) # random indices
mtbe[ind,]
```

#### Extra Questions
```{r}
#Remove the lines that contain NA values
mtbe_full = na.omit(mtbe)

#Calculate the standard deviation of the depth of wells which have “Bedrock” as the Aquifier
depth = mtbe_full[mtbe_full$Aquifier=="Bedrock",]$Depth
sd(depth)
```

## Question 6: 1.16: Earthquakes
```{r}
#Read in the earthquake file
eq = read.csv("EARTHQUAKE.csv")

#Take a random sample of thirty aftershocks
after_ind = sample(1:dim(eq)[1], 30, replace=FALSE) #Automatically pull the dimension of eq (no hard code)

aftershocks = eq[after_ind,] #Take sample of aftershocks

print(aftershocks)
```

#### Supplimentary Question 


